Analysis and Figures for Gallego et al, 2020

### 2020-08-13 Scripts/All.figures.that.go.in.the.MS.no.seasons.Rmd
All ready to go

This Markdown will make the part of the analysis that goes in the Manuscript:

```{r libraries and custom functions, echo = F, message=F}
library (RColorBrewer)
library (patchwork)
library (rstanarm)
library (fitdistrplus)
library (ggridges)
library (ggforce)
library (tidyverse)
library (cowplot)
library (concaveman)
library (vegan)
library (lubridate)
library (ggrepel)
library (factoextra)
library (nnet)
library (here)
library (conflicted)
conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")
conflict_prefer("area", "patchwork")
#functions for future Salish Sea
models <- readRDS(here("Input","full_model_output_rpk_TpH.standardized_nonSeasonal2020-06-24_3262.RDS"))

@@

stds_to_value <- read_rds(here("Input", "std_to_real.rds"))

@@

```{r }
data.plankton <- read_csv(here("Input","Combined_Biol_Env_Plankton.csv"))

@@

env.data <- read_csv(here("Input","env.data.for.log.regression.csv"))

#quality-control data
    env.data %>%
      filter(pH > 7) %>%
      mutate(
        Season = case_when(
          str_detect(sample, "1703|1710|1711|1801|1803") ~ "Winter",
          TRUE ~ "Summer"
        ),
        Area = case_when(
          str_detect(sample, "CP|LK|FH") ~ "SJI",
          TRUE ~ "Hood Canal"
        )
      ) -> now.env.data

now.env.data %>%
  gather(-Season, -sample, -Area,  key = "Predictor", value = "value") %>%
  group_by(Area, Predictor) %>%   #RPK: this was missing, which messed up the std calculations 
  mutate(std = vegan::decostand(value, method = "standardize")) -> now.env.data.with.std

#### TODO: remove

# Dataset to go between standarized and real values
now.env.data.with.std %>%
  group_by(Area, Predictor) %>%
  # summarise(mean = mean(value),
  #           sd = sd(value)) %>%
  # group_by(Area, Predictor) %>%
  nest() %>% 
  summarise(tovalue = map(data, ~ lm(value ~ std, data = .x) ),
            tostd   = map(data, ~ lm(std ~ value, data = .x)))-> transforming.df

transforming.df[1,4] %>% pull -> test
broom::tidy(test[[1]])
extract <- function (lm) {broom::tidy(lm)$estimate}

extract(test[[1]])

transforming.df %>% 
  mutate(to.values = map(tovalue, extract),
         to.std= map(tostd, extract)) ->transforming.df

@@

```{r Future scenarios}
Year.area.dataset <- function(Area, Year, n){

    Temp.conv <- stds_to_value[which (stds_to_value$Area == Area & stds_to_value$key == "Temperature"),"toValue"] %>% pull %>% .[[1]]
    pH.conv   <- stds_to_value[which (stds_to_value$Area == Area & stds_to_value$key == "pH")         ,"toValue"] %>% pull %>% .[[1]]

    one<-  data.frame(
              Area = Area,
              Year = Year,
              Temperature = r_Temperature_year_STD(Year, Area, n, 2017),
              pH          = r_pH_year_STD(Year, Area, n, 2017)
              )

    one$TemperatureCelsius = sapply(one$Temperature,function(x){Temp.conv$coefficients[1]+Temp.conv$coefficients[2]*x})
    one$pH_units            = sapply(one$pH,function(x){pH.conv$coefficients[1]+pH.conv$coefficients[2]*x})

    return(one)

}

j <- Year.area.dataset("Hood Canal", 2095, 5000)
k <- Year.area.dataset("Hood Canal", 2017, 5000)
l <- Year.area.dataset("SJI", 2095, 5000)
m <- Year.area.dataset("SJI", 2017, 5000)

simulated.scenario <- rbind(j,k,l,m)

rm(j,k,l,m)

simulated.scenario %>%
  group_by(Year, Area) %>% 
  nest() %>% 
  mutate(
    values = map (data, function (.x) {
      list(temp = fitdistr(.x$TemperatureCelsius ,
             densfun = "normal"),
           ph = fitdistr(.x$pH_units ,
             densfun = "normal"))
            }
          ),

    data = map2(data,values,  function (.x, .y){
      .x %>%
        mutate(pT = pnorm(TemperatureCelsius, mean  = .y[["temp"]]$estimate[1], sd = .y[["temp"]]$estimate[2]),
               ppH =pnorm(pH_units, mean  = .y[["ph"]]$estimate[1], sd = .y[["ph"]]$estimate[2]) )
      })
    ) %>% 
  unnest(data) %>% 
  mutate(keep = case_when(pT < 0.05 | ppH < 0.05 | pT > 0.95 | ppH > 0.95 ~ "Discard",
                          TRUE                   ~ "keep")) %>% 
  filter (keep == "keep")-> simData.nocrazys

simData.nocrazys %>% 
  select(-values) %>% write_csv(here("Output","simulation.from.std.functions.cleaned.csv"))
```

@@

simData.nocrazys %>%

saveRDS( file = here("Output","all.sim.data_climateScenarios_no.outliers.RDS"))

simData.nocrazys %>% 
  group_by( Area, Year) %>% 
  summarise_at(c("TemperatureCelsius", "pH_units"), .funs = list(min = min, max = max))

@@

We have the datasets and some plots already made -

@@

```
### Panel B Environmental variation

@@

```{r}
now.env.data %>% 
  select(TA = mean.TA, DIC = mean.DIC, everything(), -contains("%")) %>% 
  filter(!is.na(TA)) %>% 
  separate(sample, into= c("Site","Date"), remove = F )-> env.data.no.outliers

env.data.no.outliers %>% 
  pivot_longer(cols = c(-Area, -Site, -Date, -Season, -sample), names_to =   "Predictor", values_to  = "value") -> env.data.to.keep

env.data.to.keep %>% 
  mutate(Site = fct_relevel(Site, "SA", "TR", "LL","PO", "TW"),
         Predictor = fct_relevel(Predictor, "DIC", "TA", "pH")) -> env.data.to.keep

env.data.to.keep %>% 
  mutate(Area = case_when (Area == "SJI" ~ "San Juan Island",
                           TRUE          ~ Area)) %>% 
  mutate(Month = month(ymd(paste0(Date, "01")), label = T, abbr = F)) %>% 
  filter (Predictor %in% c("DIC", "TA", "pH", "Temperature", "Salinity")) %>% 
  ggplot(aes(x = `value`, y = `Month`, fill = Area)) +
  geom_density_ridges_gradient() +
 # scale_fill_viridis(name = "Temp. [F]", option = "C") +
  facet_wrap(~fct_relevel(Predictor, "DIC", "TA", "pH", "Temperature", "Salinity"), scales = "free", nrow = 1) +
  theme(legend.direction ="horizontal") -> ridges

legend <- get_legend(
  # create some space to the left of the legend
  ridges + theme(legend.box.margin = margin(0, 0, 0, 12))
)
 legend
ridges + guides(fill = "none") -> ridges
```

### Panel c: Richness variation

```{r}
boxplot.richness <- readRDS(here("Input","richness.rds"))

boxplot.richness +
  ylim (0,60) -> boxplot.richness
```
### All together

THe top will be the environmental gradients, the lower panels will have the map and the boxplot
```{r}
plot_grid(plotlist = list(map, legend, boxplot.richness),
                     align = "v",
                     ncol = 3, 
          axis = "t",
          rel_widths = c(1,0.5,0.6))   -> bottom.half.plot1

plot_grid(plotlist = list(ridges, bottom.half.plot1),
          align = "hv",
          ncol = 1, 
          nrow = 2,
          rel_heights = c(1,0.8),
          rel_widths = 1)

plot_grid(plotlist = list(legend, boxplot.richness),
                     align = "hv",
                     ncol = 1, 
          axis = "t",
          rel_heights  = c(.5,1),
          labels = c(NA,"B" )) -> bottom.right

plot_grid (plotlist = list(map, bottom.right),
           align = "hv",
           nrow = 1,
           rel_heights = 1,
           rel_widths = c(.6,0.4),
           labels = c("A", NA)) -> bottom.plot

plot_grid(bottom.plot, ridges,
          align = "h",
          nrow = 2,
          rel_heights = c(0.6, 0.4),
          labels = c(NA,"C")) +
  ggsave(filename = here("Figures","Manuscript","Fig1.png"), width = 14, height = 14, dpi = "retina")

@@

data.plankton %>% 
  separate (event, into = c("Site", "Date"), sep = "_") %>% 
  mutate(Season = case_when(str_detect(Date,"17-03|17-10|17-11|18-01|18-03") ~ "Winter",
                             TRUE ~                                         "Summer"),
         Area = case_when(Site %in% c("CP","LK", "FH") ~ "San Juan Island",
                          TRUE                         ~ "Hood Canal")) %>% 
   unite (Site, Date, col = "event", sep = "_") %>% 
  cap.functions.only3things(Area) -> joined.dataset.for.CAP
```

Step by step
```{r CAP1}
joined.dataset.for.CAP %>% 
  mutate(
    CAP.analysys.raw = map2(comm.matrix, env,

@@

So the best numbers of clusters are

@@

Now create the plot that combines Constrained Analysis of Principal Coordinates and clusterings
### Panel A: CAP analysis with environmnental correlation and cluster info
```{r Hand made CAP}

@@

  all.sim.data %>% 
    mutate (Area = fct_recode(Area, "San Juan Island" = 
                                "SJI")) %>% 
    group_by(Area) %>% 
    nest() %>% 
    mutate(pH.range = map(data, ~ seq(round(min(.x$pH_units),1), 
                                      round(max(.x$pH_units),1),
                                      length.out = 12)),
           T.range =  map(data, ~ seq(round(min(.x$TemperatureCelsius),1), 
                                      round(max(.x$TemperatureCelsius),1),
                                      length.out = 22)),
           output.range = map2(pH.range, T.range, ~ expand_grid(pH_new = .x, Temperature = .y))) %>%

    left_join(clusters) %>% 
    mutate(whichComm.res = map2(output.range,multi.mod,~ .x %>% 
                                  mutate(comm = as.numeric (apply(predict(.y, type = "probs", newdata = .x),1, which.max)) )
              ) ,
           plots = map(whichComm.res, ~ 
                         ggplot(.x, aes(y = pH_new,
                                        x = Temperature,
                                        fill = as.factor(comm),
                                        )) +
                         geom_tile()+
                         geom_label(data = . %>% group_by(comm) %>% 
                                      summarise_all( mean),
                                    aes(x= Temperature,
                                        y= pH_new,
                                        label = comm),
                                    fill = "white")+
                         scale_fill_brewer(name = "Cluster", palette = "Set2") +
                         xlab("Temperature") + ylab("pH")+
                         theme_light())) -> testquick

@@

## Get the data for all seasons and areas
 SIMPER.clusters %>%
   pull(summary.simper)-> list.of.simpers
names(list.of.simpers) <- paste(SIMPER.clusters$Area, SIMPER.clusters$Season, sep = "_")

@@

    theme_light() + theme(axis.text.y = element_text(face = "bold.italic", colour ="#3B4252", hjust = 0 ),
                          strip.background = element_rect(fill = "#3B4252"),
                          strip.text = element_text(face = "bold", color = "#EBCB8B"))} 
  ))   -> right_most

@@

We use the climate projection to predict changes in suitability

all.sims.for.projection <- simData.nocrazys %>% 
  select(Area, Year, Temperature, TemperatureCelsius, pH, pH_units)

## Get 
models %>% 
  transmute(taxa, 
            post.draws = map2 (model, taxa, ~ posterior_predict(object = .x,
                                                                newdata = all.sims.for.projection,
                                                                draws = 100) %>% 
                                 as_tibble() %>% 
                                 t(.) %>% 
                                 as_tibble() %>% 
                                 rownames_to_column("sim.row") )) -> posterior.draws.all

@@

posterior.draws.all %>% 
 left_join(phylum.info %>% dplyr::select(phylum, taxa)) %>% 
  group_by(phylum) %>% # Add phylum information
  nest() %>% 
   mutate (data = map (data, ~ .x %>% 
                        unnest() %>% 
                        pivot_longer(cols = c(-taxa, -sim.row),
                                    names_to = "draw.row", values_to = "presence") %>% 
  group_by(taxa, sim.row) %>% 
  summarise(prob.of.presence = mean(presence)) %>% 
  left_join(all.sims.for.projection %>%  
              rownames_to_column("sim.row"))

  )) -> posterior.probabilities.by.taxa
```
Now we have a nested dataframe per each phylum. Not that we need it. But the nested part is a dataframe with the mean result of the 100 draws for each environmental point.

### Richness by phyla with time
 First create the square matrix of points - the ggplot geometries requires one entry per grid - so we have to comply
```{r create full grid}

all.sims.for.projection  %>% 
  group_by(  Area) %>% 
  # filter (keep == "keep") %>%   
  summarise_at(c("TemperatureCelsius", "pH_units"), list(max = max, min = min)) %>%

  group_by ( Area) %>% 
  nest() %>% 
  mutate(data = map (data, ~ expand_grid(TemperatureCelsius = seq(round(.x$TemperatureCelsius_min,1), round(.x$TemperatureCelsius_max, 1),by = 0.1),
                                         pH_units = seq(round(.x$pH_units_min,1), round(.x$pH_units_max,1), by = 0.1)))) %>% 
  unnest(data) -> full.grid
```

```{r add standarized temperatures to the grid}

full.grid %>% 
  ungroup() %>% 
  rownames_to_column("Case") %>% 
  rename(Temperature = TemperatureCelsius,
         pH = pH_units) %>%

  pivot_longer(cols = c(Temperature, pH),
               names_to  = "Predictor",
               values_to = "value") %>% 
  group_by(Area, Predictor) %>% 
  nest() %>% 
  left_join(transforming.df) %>% 
  mutate(data = map2(data, to.std, ~.x %>% 
                       mutate(std = .y[1]+.y[2]*value ))) %>% 
  select(Area, Predictor, data) %>% 
  unnest(data) %>% 
  pivot_wider(names_from = Predictor,
              values_from = c(std,value)) %>% 
  rename(Temperature = std_Temperature,
         pH = std_pH,
         TemperatureCelsius = value_Temperature,
         pH_units           = value_pH) -> full.grid
```

`full.grid` has the temperature and pH values we want to project our models to - so now we use it with the `models` output.

```{r calculate the 100 draws from the posterior}
models %>% 
  ungroup() %>%  #rpk added
  transmute(taxa, 
            post.draws = map2 (model, taxa, ~ posterior_predict(object = .x,
                                                                newdata = full.grid,
                                                                draws = 100) %>% 
                                 as_tibble() %>% 
                                 t(.) %>% 
                                 as_tibble() %>% 
                                 rownames_to_column("sim.row") )) -> posterior.draws.raster
```

```{r calculate richness}
posterior.draws.raster %>% 
 left_join(phylum.info %>% dplyr::select(phylum, taxa)) %>% 
  group_by(phylum) %>% 
  nest() %>% 
  mutate(data = map (data, ~ .x %>% unnest(cols= post.draws) %>% 
                       pivot_longer(cols = c(-taxa, -sim.row),
                                    names_to = "draw.row", values_to = "presence") %>% 
                       group_by(draw.row, sim.row) %>% 
                       summarise(tot = sum(presence)) %>% 
                       group_by(sim.row) %>% 
                       summarise_at(.vars = "tot", .funs = list(Richness = mean,sd = sd)))) %>% 
  unnest(data) -> ready_plot

ready_plot %>% 
  left_join(full.grid %>%  
              rownames_to_column("sim.row")) -> ready_plot

# Normailzed the data for each phyla

ready_plot %>% 
  group_by(phylum) %>% 
  mutate(st.richness = Richness / max(Richness)) -> ready_plot

@@

```{r plot3 function}
plot3 <- function(tibble,string , response) {

@@

  tibble %>% 
    ungroup() %>% 
    mutate(Area = as.character(Area)) %>% 
    mutate(Area = case_when(Area == "SJI" ~ "San Juan Island",
                            TRUE          ~ Area)) %>% 
    group_by(TemperatureCelsius, pH_units, Area) %>%

@@

    ggplot(aes(x = round(TemperatureCelsius,10),
             y = round(pH_units, 10))) +

@@

  # geom_density2d(data = simData %>%
  #                   filter (Year %in% c(2015, 2095)),
  #                 aes(x= TemperatureCelsius, y = pH_units, group= Year, color = as.factor(Year)), bins = 5, alpha = 0.5) +

>>>>

This Markdown will make the part of the analysis that goes in the Manuscript: The starting point are the Abundance Table (number of reads per ASV per sample) and the key file (the ASV file has the Hash of the sequences, and the key has the actual DNA sequence). 
If you want to process the raw fastq files to obtain these two files, you have to use the pipeline available in http::/github.com/ramongallego/demultiplexer_for_dada2

```{r libraries and custom functions, echo = F, message=F}
library (RColorBrewer)
library (patchwork)
library (rstanarm)
library (fitdistrplus)
library (ggridges)
library (ggforce)
library (tidyverse)
library (concaveman)
library (vegan)
library (lubridate)
library (ggrepel)
library (factoextra) # Needed for visualizing clusters in Fig2
library (nnet) # Needed for multinomial regression in Fig 2
library (here)
library (conflicted)

conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")
conflict_prefer("area", "patchwork")
#functions for future Salish Sea
models <- readRDS(here("Input","finalModel_20200807.RDS"))

@@

@@

```{r loading data}
data.plankton <- read_csv(here("Input","Combined_Biol_Env_Plankton.csv"))

@@

# How to go from real data from standardized data and back
value2std <- function(df){lm(std~value, data = df)}
std2value <- function(df){lm(value~std, data = df)}

stds_to_value <- data.plankton %>% 
  select(sample = event, Area, mean.TA, mean.DIC = mean.DIC.new.sal,Salinity = Salinity.new ,Temperature, Omega.aragonite, pH = pH_new ) %>% 
  distinct() %>%  
  filter(pH > 7.4) %>% 
  ungroup() %>% 
  pivot_longer(cols = c(-sample, -Area),
               names_to = "key",
               values_to = "value") %>% 
          group_by(key, Area) %>%   #calculating different scale for each area; standardizing to different means
          mutate(std = vegan::decostand(value, method = "standardize")) %>%  
          ungroup() %>% 
          mutate(Area = as.factor(Area)) %>% 
          filter(key %in% c("Temperature", "pH"))  %>% 
          nest(dataset = c(sample, value, std)) %>% 
          mutate(toStd = map(dataset, value2std),
                 toValue = map(dataset, std2value))

@@

```{r Future scenarios}
future.scenario <- tibble(Area = rep(c("Hood Canal", "SJI"), 6),
                  Year = rep(c(2017, 2091:2095), each = 2))  %>% 
  mutate(Temperature = map2(Area, Year, ~  r_Temperature_year(area = .x,year =  .y, 1000)),
         pH = map2(Area, Year, ~  r_pH_year(area = .x,year =  .y, 1000)))
future.scenario %>% unnest(cols = c(Temperature, pH)) -> future.scenario

# Make 2095 the 95% interval of the 2091-2095 scenarios 
future.scenario %>% 
  mutate(Year = case_when(Year == 2017 ~ 2017, 
                          TRUE        ~ 2095)) %>%
  group_by(Year, Area) %>% 
  nest() %>% 
  mutate(values = map (data, function (.x) {
                              list(temp = fitdistr(.x$Temperature ,
                                                 densfun = "normal"),
                                   ph = fitdistr(.x$pH ,
                                                 densfun = "normal")) 
            }),
         data = map2(data,values,  function (.x, .y){
                              .x %>%
                                mutate(pT = pnorm(Temperature, 
                                                  mean  = .y[["temp"]]$estimate[1],
                                                  sd = .y[["temp"]]$estimate[2]),
                                       ppH =pnorm(pH,
                                                  mean  = .y[["ph"]]$estimate[1], 
                                                  sd = .y[["ph"]]$estimate[2]) )
           })) %>%

  unnest(data) %>%

  mutate(keep = case_when(pT < 0.03 | ppH < 0.03 | pT > 0.98 | ppH > 0.98 ~ "Discard",
                          TRUE                   ~ "keep")) %>% 
  filter (keep == "keep") -> future.scenario

future.scenario %>% 
  select(-values) %>% write_csv(here("Output","simulation.from.std.functions.cleaned.csv"))
```

We are going to use this future simulation in two ways:

  * As it is, to explore the changes in suitability for each taxa between present and future sceanrios
  * As a means to establish a continuous grid of Temperature and pH, so we can visualize how broader changes might occur.

```{r create the grid}

future.scenario %>% 
  group_by(  Area) %>% 
  # filter (keep == "keep") %>%   
  summarise_at(c("Temperature", "pH"), list(max = max, min = min)) %>%

  group_by ( Area) %>% 
  nest() %>% 
  mutate(data = map (data, ~ expand_grid(Temperature = seq(round(.x$Temperature_min,1), round(.x$Temperature_max, 1),by = 0.1),
                                         pH = seq(round(.x$pH_min,1), round(.x$pH_max,1), by = 0.1)))) %>% 
  unnest(data) -> full.grid
```

# Add the standardized values for each dataset

Because the standarization was done for each region separately, we have to convert them into standard values seapratedly as well.

```{r add standard values to both ds}

convertT <- function( values, area){
  stds_to_value %>% 
    filter (Area == area, key == "Temperature") %>% 
    pull(toStd) -> mod
    predict(mod[[1]],newdata = data.frame (value = values))
}
convertpH <- function( values, area){
  stds_to_value %>% 
    filter (Area == area, key == "pH") %>% 
    pull(toStd) -> mod
    predict(mod[[1]],newdata = data.frame (value = values))
}

full.grid %>% 
  rename(TemperatureCelsius = Temperature,
         pH_units           = pH) %>% 
  group_by(Area) %>% 
  nest() %>% 
  mutate(data =  map2(Area, data, ~ .y %>% 
                            mutate(Temp.std =convertT( area = .x, values =.y$TemperatureCelsius),
                                   pH.std   =convertpH(area = .x, values =.y$pH_units) ))) %>% 
  unnest(data) -> full.grid

future.scenario %>% 
  select(Area, Year, TemperatureCelsius = Temperature, pH_units = pH) %>% 
  group_by(Area) %>% 
  nest() %>% 
  mutate(data =  map2(Area, data, ~ .y %>% 
                            mutate(Temp.std =convertT( area = .x, values =.y$TemperatureCelsius),
                                   pH.std   =convertpH(area = .x, values =.y$pH_units) ))) %>% 
  unnest(data) -> future.scenario
```

@@

full.grid %>%

saveRDS( file = here("Output","Temp.pH.grid.RDS"))

future.scenario %>% 
  write_rds(file = here("Output", "future.simulation.rds"))

future.scenario %>% 
  group_by( Area, Year) %>% 
  summarise_at(c("TemperatureCelsius", "pH_units"), .funs = list(min = min, max = max))

@@

We have the datasets and the map already made

@@

```
### Panel B  Richness variation

```{r Species richness}

data.plankton %>% 
  left_join(phylum.info) %>% 
  group_by(event, phylum) %>% 
  summarise (Richness = n_distinct(taxa)) %>% 
  group_by(event) %>% 
  mutate(Total.Richness = sum(Richness)) %>% 
  separate(event, into = c("Site", "Date"), sep = "_") %>% 
  mutate(Area = case_when(Site %in% c("CP", "LK", "FH") ~ "San Juan Island",
                          TRUE                          ~ "Hood Canal"),
         Site = fct_relevel(Site, "CP", "LK", "FH","SA", "TR", "LL","PO", "TW")) -> data.for.richness

data.for.richness %>%   
ggplot(aes(x = Site, y= Total.Richness )) +
  geom_boxplot(aes(fill = Area)) +
  lims (y= c(0,max(data.for.richness$Total.Richness))) +
  labs (y = "Species richness") -> left.rich.plot

data.for.richness %>% 
  filter (phylum %in% c("Bacillariophyta", "Dinophyceae")) %>% 
  ggplot(aes(x = Site, y= Richness )) +
  geom_boxplot(aes(fill = Area)) +
  facet_wrap(~phylum, nrow = 2) +
  labs (y = "") -> right.rich.plot

left.rich.plot + right.rich.plot + plot_layout(tag_level = "new") -> boxplot.richness
```

### Panel C Environmental variation

@@

```{r envRidges}
data.plankton %>% 
  select(sample = event, Area, TA = mean.TA, DIC = mean.DIC.new.sal,Salinity = Salinity.new ,Temperature, Omega.aragonite, pH = pH_new ) %>% 
  distinct() %>%  
  filter(pH > 7.4) %>% 
  ungroup() %>% 
  filter(!is.na(TA)) %>% 
  separate(sample, into= c("Site","Date"), remove = F, sep = "_" ) %>% 
  pivot_longer(cols = c(-Area, -Site, -Date,  -sample), names_to =   "Predictor", values_to  = "value")  %>% 
  mutate(Site = fct_relevel(Site, "SA", "TR", "LL","PO", "TW"),
         Predictor = fct_relevel(Predictor, "DIC", "TA", "pH")) %>% 
  mutate(Month = month(ymd(Date), label = T, abbr = F)) %>% 
  filter (Predictor %in% c("DIC", "TA", "pH", "Temperature", "Salinity")) %>% 
  ggplot(aes(x = `value`, y = `Month`, fill = Area)) +
  geom_density_ridges_gradient() +
 # scale_fill_viridis(name = "Temp. [F]", option = "C") +
  facet_wrap(~fct_relevel(Predictor, "DIC", "TA", "pH", "Temperature", "Salinity"), scales = "free", nrow = 1) +
  theme(legend.direction ="horizontal", axis.title.x = element_blank()) -> ridges

ridges + guides(fill = "none") -> ridges
```

### All together

THe top will be the environmental gradients, the lower panels will have the map and the boxplot. Create a custom palette of colors for the two regions

```{r assmbling Fig1}
Site.palette <- c("#d8b365", "#5ab4ac")

layout <- "
AAABBB
AAABBB
AAACCC
DDDDDD
DDDDDD"

wrap_plots(A = map ,
           B = boxplot.richness,
           C = guide_area(),
           D = ridges + labs(x = ""),
            design = layout,
            heights = c(10,10,1,10,10),
           guides = "collect")  +
  plot_annotation(tag_levels = c("A", "1"), tag_sep = ".")  & 
  theme_minimal() &
  theme(legend.direction = "horizontal",
        strip.background = element_rect(fill = "#2E3440"),
        strip.text = element_text(face = "bold", color = "#D8DEE9") ) &
  scale_fill_manual(values = Site.palette) -> Fig1

Fig1
ggsave(Fig1, filename = here("Figures","Manuscript","Fig1.png"), width = 14, height = 12, dpi = "retina")

@@

data.plankton %>% 
  separate (event, into = c("Site", "Date"), sep = "_") %>% 
  mutate(Area = case_when(Site %in% c("CP","LK", "FH") ~ "San Juan Island",
                          TRUE                         ~ "Hood Canal")) %>% 
   unite (Site, Date, col = "event", sep = "_") %>% 
  cap.functions.only3things(Area) -> joined.dataset.for.CAP
```

Step by step. Create a column for each stage of the CAP analysis - and stop to check the signifficance of the CAP
```{r CAP1}
joined.dataset.for.CAP %>% 
  mutate(
    CAP.analysys.raw = map2(comm.matrix, env,

@@

Now we want to see if there are clusters associated with this. A first step is to choose how many clusters do we need to form. There are several methods for evaluating the goodness of fit of the number of clusters. 
So the best numbers of clusters are

@@

Now create the plot that combines Constrained Analysis of Principal Coordinates and clusterings
### Panel A: CAP analysis with environmnental correlation and cluster info

```{r Hand made CAP}

@@

full.grid %>% 
  mutate (Area = str_replace(Area, replacement = "San Juan Island", pattern =  "SJI")) %>%
  rename(Temperature = TemperatureCelsius,
         pH_new = pH_units) %>% 
    group_by(Area) %>% 
    nest() %>% 
    left_join(clusters) %>% 
    mutate(whichComm.res = map2(data,multi.mod,~ .x %>% 
                                  mutate(comm = as.numeric (apply(predict(.y, type = "probs", newdata = .x),1, which.max)) )
              ) ,
           plots = map(whichComm.res, ~ 
                         ggplot(.x, aes(y = pH_new,
                                        x = Temperature,
                                        fill = as.factor(comm),
                                        )) +
                         geom_tile()+
                         geom_label(data = . %>% group_by(comm) %>% 
                                      summarise_all( mean),
                                    aes(x= Temperature,
                                        y= pH_new,
                                        label = comm),
                                    fill = "white")+
                         scale_fill_brewer(name = "Cluster", palette = "Set2") +
                         xlab("Temperature") + ylab("pH")+
                         theme_light())) -> testquick

@@

## Get the data for all  areas
 SIMPER.clusters %>%
   pull(summary.simper)-> list.of.simpers
names(list.of.simpers) <- paste(SIMPER.clusters$Area, SIMPER.clusters$Season, sep = "_")

@@

    theme_light() + theme(axis.text.y = element_text(face = "bold.italic", colour ="#3B4252", hjust = 0 ),
                          strip.background = element_rect(fill = "#2E3440"),
        strip.text = element_text(face = "bold", color = "#D8DEE9")
        # strip.background = element_rect(fill = "#3B4252"),
        # strip.text = element_text(face = "bold", color = "#EBCB8B")
        )} 
  ))   -> right_most

@@

We use the climate projection to predict changes in suitability. First panel is a visualization of the suitability of planktonic taxa, highlighting those species with a more pronounced change.

```{r Prepare dataset}
models$data %>% 
  distinct(Area, taxa, taxonAreaIndex) -> keepers

expanded.future <- future.scenario %>% 
  select(Area, Year, Temperature = Temp.std, TemperatureCelsius, pH = pH.std, pH_units)

## Get

expand_grid(models$data %>% distinct(taxa), expanded.future) -> expanded.future

left_join(expanded.future, keepers) -> expanded.future
```

```{r Calculate the 100 draws}
posterior_predict(object = models,
                   newdata = expanded.future,
                   draws = 100) -> posterior.predict.all.taxa

@@

expanded.future %>% left_join(phylum.info) -> expanded.future

expanded.future$present <- colMeans(posterior.predict.all.taxa)

expanded.future$sd <- apply(posterior.predict.all.taxa,2,sd )
```

Now choose for each phyla, the taxa with more dramatic change

```{r}
expanded.future %>% 
  left_join(phylum.info) %>% 
  nest(-phylum) %>% 
  # ungroup() %>% 
  # filter (phylum %in%c("Dinophyceae", "Bacillariophyta")) %>% 
  mutate(suma = map (data,~.x %>% group_by(taxa,Area, Year) %>%
                       summarise(tot = sum(present)) %>%
                       ungroup() %>% 
                       pivot_wider(names_from = Year, values_from = tot) %>% 
                        group_by(taxa) %>% mutate(diff = `2017` - `2095`)
                     )) %>% 
  select(phylum, suma) %>% 
  unnest(suma) %>%
  group_by(phylum) %>% arrange(desc(abs(diff))) %>% 
  separate(taxa, into = c("family","genus", "species"), sep = "\\|", remove = F) %>%
  mutate(label = case_when((species == "NA" & genus == "NA") ~ paste0(family, " sp."),
                           species == "NA" ~  paste0(genus, " sp."),
                           TRUE            ~ species)) -> taxa.and.diff

```{r Fig3left}

expanded.future %>% 
  filter(taxa %in% ahorasi) %>%
  separate(taxa, into = c("family","genus", "species"), sep = "\\|", remove = F) %>%
  mutate(label = case_when((species == "NA" & genus == "NA") ~ paste0(family, " sp."),
                           species == "NA" ~  paste0(genus, " sp."),
                           TRUE            ~ species))    %>%
  mutate(Area = as.character(Area)) %>%
  mutate(Area = case_when(Area == "SJI" ~ "San Juan Island",
                          TRUE          ~ Area)) %>%
  mutate(Year = fct_rev(as.factor(Year) )) %>% 
  nest(-phylum) -> input.for.left

input.for.left %>% 
  mutate( plot.by.taxa = map2 (data, phylum, ~ .x %>%
                                 ggplot(aes(x= present,
                                            y = fct_relevel(label,reorder.labels))) +
                                 scale_fill_brewer(name = "Year",  type = "qual", palette = 4, direction = 1,
                                                   breaks=c("2017","2095")) +
                                 labs (x = "Probability of presence",
                                       y = "") +
                                 facet_wrap(~Area) +
                                 theme_minimal() +
                                 theme(strip.text.y = element_text(angle = 0),
                                       axis.text.y = element_text(face = "italic"),
                                       legend.direction = "horizontal",
                                       legend.position  = "bottom",
                                       plot.title.position = "plot",
                                       plot.title = element_text(face = "bold",vjust =  -5 ),
                                       axis.title.x.top  = element_text(face = "bold",vjust = 3) ) +
                                 ggtitle(.y)  +
                                 scale_x_continuous(limits = c(0,1),position = "top") )) %>%
  pull(plot.by.taxa) %>% 
  set_names(input.for.left$phylum) -> fig3.left

fig3.left["Dinophyceae"]
```

### Richness by phyla with time
 First prepare the grid - expand it to include the taxa
```{r}
full.grid %>% 
  select(Area,  Temperature = Temp.std, TemperatureCelsius, pH = pH.std, pH_units) -> full.grid

expand_grid(models$data %>% distinct(taxa), full.grid) -> full.grid

left_join(full.grid, keepers) -> full.grid
```

Now calculate the posterior from each point in the grid
```{r}
posterior_predict(object = models,
                   newdata = full.grid,
                   draws = 100) -> posterior.predict.all.taxa

@@

Add the phylum information and calculate the richness at each point
```{r}
full.grid %>% left_join(phylum.info) -> full.grid

full.grid$present <- colMeans(posterior.predict.all.taxa)

full.grid$sd <- apply(posterior.predict.all.taxa,2,sd )

full.grid %>% 
  group_by(Area, Temperature = TemperatureCelsius, pH = pH_units, phylum) %>% 
  summarise(richness = sum(present),
            sd       = sum(sd^2)) %>% 
#  left_join(richness.by.phylum) %>% 
  group_by(phylum) %>% 
  mutate(st.richness = richness/max(richness)) -> richness.by.phylum
```

```{r plot4 function}
plot4 <- function(tibble, response) {

@@

  tibble %>% 
    ungroup() %>% 
    mutate(Area = as.character(Area)) %>% 
    mutate(Area = case_when(Area == "SJI" ~ "San Juan Island",
                            TRUE          ~ Area)) %>% 
    mutate(Temperature = round(Temperature, 1),
           pH = round(pH, 1)) %>% 
    group_by(Temperature, pH, Area) %>%

@@

    ggplot(aes(x = Temperature,
             y = pH) ) +

@@

  scale_color_brewer(name = "Year",  type = "qual", palette = 4, direction = -1, breaks=c("2017","2095")) +
  ggforce::geom_mark_hull(data = all.sims.for.projection %>%
                            ungroup() %>%
                            mutate(Area = as.character(Area)) %>% 
                            mutate(Area = case_when(Area == "SJI" ~ "San Juan Island",
                            TRUE          ~ Area)) %>%
                            filter (Year %in% c(2017, 2095)) ,
                          aes(x= TemperatureCelsius, y = pH_units, group= Year, color = as.factor(Year),label = as.factor(Year)),
                          expand = unit(0.5, "mm"),
                 label.margin = margin(1, 1, 1, 1, "mm"),
                 label.buffer = unit(0, "mm"),
                 con.cap = 0 ,
                 alpha = 0.5) +
     facet_grid(.~Area) +
  guides(color = "none") +
    labs (x = "Temperature",
          y= "pH") +
    theme_minimal()+
    theme(legend.position = "bottom",
          legend.key.width  = unit(2 ,"cm"),
          axis.title.x = element_text(face = "bold"),
          legend.direction = "horizontal") +
    scale_x_continuous(position = "top")
}
```

```{r plot4 in play}
richness.by.phylum %>% 
  nest(-phylum) %>% 
  mutate(plot = map(data, ~ plot4(.x, response = st.richness)))  -> list.of.richness.plots
